Parallelizing AdaBoost and Decision Trees

### 2015-08-21 paper.tex
paper changes

\section{Introduction} Binary and multiclass classification is a fundamental
problem in machine learning and is commonly one of the first concepts to be
studied in an introductory course. In the context of binary classification, a
labeled observation is a coordinate pair $(x, y)$ where $x\in\mathbb{R}^d$ is a
vector of data associated with the observation (feature vector) and $y\in\{1,
-1\}$ is the corresponding label of the observation. For example, $x_i$ might
be a point in $\mathbb{R}^2$ representing the height and weight of the $i$th
individual in a data set while the value of $y_i$ may correspond to the
person's gender. The goal in classification is to accurately predict the label
of a given observation using its feature vector. In this setting, the scenario
of supervised learning involves using a training set of labeled observations
$(x_i, y_i)$ to construct a prediction function $h$, one that takes a feature
vector $\hat{x}$ and gives a prediction on the corresponding label
$h(\hat{x})=\hat{y}$. It is common to use a separate set of test data to check
the accuracy of the predictor, data that hasn't been used for training. There
are many methods used to obtain a predictor for classification in supervised
learning, and the class of ensemble methods concerns combining several
predictors to obtain one.

Suppose now there exists an algorithm that, for any set of training data, is
guaranteed to give you a predictor that has at least 50\% accuracy. Such an
algorithm is called a weak learning algorithm since it yields a predictor that
performs marginally better than guessing randomly. These prediction functions
that a weak learning algorithm returns are called base classifiers, and a
natural question arises about whether we can use a combination of weak learners
to get a stronger one with much better accuracy and what this sort of ensemble
method would look like. Ensemble methods like this are called boosting, and the
algorithm AdaBoost introduced in \cite{no1} accomplishes this through
successive rounds of training a weak learner to the data and subsequently
re-weighting the training samples \cite{no1}.

@@

\section{AdaBoost} The beauty of AdaBoost is in using a relatively inaccurate
weak learning algorithm to produce a much more accurate prediction function. In
each round of AdaBoost for $t$ from 1 to $T$, the weak learner selects a
base classifier $h_t: \mathbb{R}^n\rightarrow\{1,-1\}$ from a set of functions
$H$, in particular choosing the one with the least error
$\epsilon_t=\sum_{i=1}^{n}D_t(i)I(h_t(x_i)\neq y_i)$ (Line 4) where $D_t(i)$ is
the weight associated with the $i$th sample. The algorithm then reweights the
training samples in Line 8 to place more emphasis on those that that the chosen
classifier $h_t$ incorrectly predicted, so that in future rounds the base
classifiers are tailored to correcting these mistakes. The predictor that is
returned after all rounds is a linear combination of all $h_t$'s. The
coefficient $\alpha_t$ gives more weight in the final predictor to those rounds
where the classifier performed better. $Z_t$ is a normalization constant. We
give the algorithm below.\\ \\
{\sc AdaBoost}: $S = \{(x_i,y_i):i = 1\ldots n\}$
\begin{enumerate}
\itemsep1pt \parskip0pt \parsep0pt
\item {\bf for $i=1$ to $n$}
\item \quad $D_1(i) = \frac{1}{n}$
\item {\bf for $t=1$ to $T$}
\item \quad $h_t$ = {\sc WeakLearner}$(S\sim D_t)$ with weighted error $\epsilon_t$
\item \quad $\alpha_t = \frac{1}{2}\log(\frac{1-\epsilon_t}{\epsilon_t})$
\item \quad $Z_t = 2[\epsilon_t(1-\epsilon_t)]^{1/2}$ (normalization factor)
\item \quad {\bf for $i = 1$ to $n$}
\item \quad \quad $D_{t+1}(i) = \frac{D_t(i)\exp(-\alpha_t y_i h_t(x_i))}{Z_t}$
\item $g = \sum_{t=1}{T} \alpha_t h_t$
\item {\bf return} $\mbox{sign}(g)$
\end{enumerate}

>>>>

\section{Introduction} Binary and multiclass classification is a fundamental
problem in machine learning and is commonly one of the first concepts to be
studied in an introductory course. In the context of binary classification, a
labeled observation is a coordinate pair $(x, y)$ where $x\in\mathbb{R}^d$ is a
vector of data associated with the observation (feature vector) and $y\in\{1,
-1\}$ is the corresponding label of the observation. For example, $x_i$ might
be a point in $\mathbb{R}^2$ representing the height and weight of the $i$th
individual in a data set while the value of $y_i$ may indicate the individual's
gender. The goal in classification is to accurately predict the label
of a given observation using its feature vector. In this setting, the scenario
of supervised learning involves using a training set of labeled observations
$(x_i, y_i)$ to construct a prediction function $h$, one that takes a feature
vector $\hat{x}$ and gives a prediction on the corresponding label
$h(\hat{x})=\hat{y}$. To check the accuracy of the predictor it is common to use a separate set of test data, i.e. data that hasn't been used for training. There are many methods used to obtain a predictor for classification in supervised
learning, and the class of ensemble methods concerns combining several
predictors together.

Suppose now there exists an algorithm that, for any set of training data, is
guaranteed to give you a predictor that has at least 50\% accuracy. Such an
algorithm is called a weak learning algorithm since it yields a predictor that
performs marginally better than guessing randomly. The prediction functions
returned by a weak learning algorithm are called base classifiers, and a
natural question arises asking whether a combination of weak learners can be used
to make a strong learner, or one that can achieve a certain level of accuracy.
Ensemble methods like this are called boosting, and the
algorithm AdaBoost introduced in \cite{no1} accomplishes this through
successive rounds of training a weak learner to the data and subsequently
re-weighting the training samples \cite{no1}.

@@

\section{AdaBoost} The beauty of AdaBoost is in using a relatively inaccurate
weak learning algorithm to produce a much more accurate prediction function. In
each round of AdaBoost for $t$ from 1 to $T$, the weak learner selects a
base classifier $h_t: \mathbb{R}^n\rightarrow\{1,-1\}$ from a set of functions
$H$, in particular choosing the one with the least error
$\epsilon_t=\sum_{i=1}^{n}D_t(i)I(h_t(x_i)\neq y_i)$ (Line 4) where $D_t(i)$ is
the weight associated with the $i$th sample. The algorithm then re-weights the
training samples in Line 8 to place more emphasis on those that that the chosen
classifier $h_t$ incorrectly predicted, so that in future rounds the base
classifiers are tailored to correcting these mistakes. The predictor that is
returned after all rounds is a linear combination of all $h_t$'s. The
coefficient $\alpha_t$ gives more weight in the final predictor to those rounds
where the classifier performed better. $Z_t$ is a normalization constant. We
give the algorithm below.\\ \\
{\sc AdaBoost}: $S = \{(x_i,y_i):i = 1\ldots n\}$
\begin{enumerate}
\itemsep1pt \parskip0pt \parsep0pt
\item {\bf for $i=1$ to $n$}
\item \quad $D_1(i) = \frac{1}{n}$
\item {\bf for $t=1$ to $T$}
\item \quad $h_t$ = {\sc WeakLearner}$(S\sim D_t)$ with weighted error $\epsilon_t$
\item \quad $\alpha_t = \frac{1}{2}\log(\frac{1-\epsilon_t}{\epsilon_t})$
\item \quad $Z_t = 2[\epsilon_t(1-\epsilon_t)]^{1/2}$ (normalization factor)
\item \quad {\bf for $i = 1$ to $n$}
\item \quad \quad $D_{t+1}(i) = \frac{D_t(i)\exp(-\alpha_t y_i h_t(x_i))}{Z_t}$
\item $g = \sum_{t=1}{T} \alpha_t h_t$
\item {\bf return} $\mbox{sign}(g)$
\end{enumerate}
